Accelerated evolution of convective simulations

### 2018-03-09 tex/bvp_paper.tex
adds new fig

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of systems in nature.  For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Unfortunately, convective systems generally evolve over the course of thermal
timescales, which are often much larger than
dynamical timescales. Therefore resolving dynamics in atmospheres which are sufficiently
thermally relaxed remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 10$ minutes at the solar surface, 
the Sun rotates roughly once every month)
compared to the Sun's Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale becomes intractably large
compared to dynamical timescales \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In other systems, 
the approximate state of the evolved solution can be estimated. There, a
set of initial conditions which is close to the evolved state can be 
derived analytically \cite{couston&all2017, brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
However, such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for tens to hundreds of freefall timescales \cite{stevens&all2011}, much less the
thousands of freefall times required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of non-converged
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean thermodynamic profile towards its evolved state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from evolution through a full thermal diffusion timescale. Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except where it is $\rho = \rho_0(1  - \alpha T_1)$ on the term where
the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, 
acts in the vertical momentum equation.
The equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally periodic with an extent of $x, y \in [0, \Gamma]$,
where $\Gamma = 2$ is the aspect ratio. 
In 2D simulations, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w = 0 \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. (\ref{eqn:bcs}) are the simplest
to implement in the process of accelerated evolution (see section \ref{subsection:ae})
we study here.

We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns. (\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
The linear terms (on the LHS of the equations) are solved implicitly,
while the nonlinear terms (RHS) are explicitly solved.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.

As initial conditions, we fill $T_1$ with
random white noise whose magnitude is $10^{-6}\mathcal{P}$.
This ensures that the initial perturbations are much smaller than the
evolved convective temperature perturbations, even at large Ra.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power.

@@

\subsection{The method of Accelerated Evolution}
\label{subsection:ae}
Here we describe a method of Accelerated Evolution (AE), which we use 
to rapidly evolve the thermodynamic state of convective simulations.  
We compare this AE method to Standard Evolution
(SE), in which we naively evolve the atmosphere for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. As Ra increases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
in freefall time units.
For an example of time saving achieved by using AE, we compare
energy traces at $S = 10^5$ from a SE run in Fig. \ref{fig:time_trace}a to an AE run
in Fig. \ref{fig:time_trace}c.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig. \ref{fig:time_trace}b.
At early stages in the simulation, these flux profiles are highly asymmetric,
with more flux exiting the atmosphere at the upper boundary than 
the fixed-flux lower boundary is providing.
By calculating the total flux,
F$_{\text{tot}} =$ F$_{\text{conv}} +$ F$_{\text{cond}}$, we derive the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{conv}}(z) = 1$
and $f_{\text{cond}}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\text{cond}}(z) = 1$ and $f_{\text{conv}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

\begin{figure*}[t!]
\includegraphics[width=\textwidth]{./figs/time_trace.png}
\caption{(a) Kinetic energy (black) and mean temperature (blue)  vs. time are shown
for a SE run at $S = 10^5$. The mean evolved values of kinetic energy and mean temperature,
averaged over the time shaded in green,
are denoted by the horizontal dashed lines. (b) The time- and horizontally-averaged
flux profiles are shown for the times highlighted in orange in (a).
(c) The same quantities as in (a) are shown, but for AE at the same parameters.
The axes are scaled identically in (a) and (c), and the AE method is used three times, marked by
the numbered arrows in (c). The fluxes averaged over the green shaded region of (c)
are shown in (d). The difference between
the fluxes in the AE and SE solutions is shown in (e). The dashed
vertical line in (a) and (c) denotes the simulation time at which evolved measurements
are started for the AE case.
\label{fig:time_trace} }
\end{figure*}

In a time-stationary state, the horizontal- and time-average of
Eqns. (\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}), neglecting terms which
vanish due to symmetry, are
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{conv, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, 
statistically stationary state. Furthermore, under the specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$,
\emph{the mean thermodynamic structure of the system is fully specified}.

Thus, the AE method is simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the fraction,
as a function of height, by which the flux in the system must be reduced to go from
an initial state (Fig. \ref{fig:time_trace}b) to an evolved state (Fig. \ref{fig:time_trace}d).
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}_{x,y}$, by $\sqrt{\xi}$, such 
that the advected, flux-carrying perturbations in the temperature are appropriately
diminished.  We then solve a 1D boundary value problem including
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}) using F$_{\text{conv, ev}}$
and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$, after diminishing it by a factor
of $\xi$. We use mixed thermal boundary conditions in this study, 
as in Eqn. (\ref{eqn:bcs}),
because $F_{\text{bot}}$ is exactly specified, and the temperature profile is fixed 
at a reference value for the AE solve.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

The AE method converges atmospheres from extreme flux disequilibrium states 
(Fig. \ref{fig:time_trace}b) on short dynamical timescales (Fig. \ref{fig:time_trace}c)
to a converged state (Fig. \ref{fig:time_trace}d) whose flux balance is very
close to the balance present in the corresponding SE solution (Fig. \ref{fig:time_trace}e).
While the application of the AE method advances solutions toward the proper solution,
we find that multiple applications of the method is the best means of achieving converged
solutions quickly.

%%%%%%%%%%%%
%%%%%%%%%%%
% RESULTS
%%%%%%%%%%%
%%%%%%%%%%%%
\vspace{-0.5cm}
\section{Results}
\label{sec:results}
We study evolved SE solutions in $S \in (1, 10^5]$ in 2D and $S \in (1, 10^4]$ in
3D. We compare their properties to
accelerated evolution (AE) runs in $S \in (1, 10^7]$ in 2D and
$S \in (1, 10^4]$ in 3D.
We refer the reader to appendix \ref{appendix:run_table} for a full list of
simulations.

>>>>

\section{Introduction}
\label{sec:intro}
Astrophysical convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of systems in nature.  For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use fully implicit 
\cite{viallet&all2011, viallet&all2013, viallet&all2016} or mixed
implicit-explicit \cite{lecoanet&all2014, anders&brown2017, bordwell&all2018} 
timestepping mechanisms have made it feasible to study
convection at low Mach numbers, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Unfortunately, convective systems generally evolve over the course of thermal
timescales, which are often much larger than all relevant
dynamical timescales. Therefore resolving dynamics in atmospheres which are sufficiently
thermally relaxed remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 10$ minutes at the solar surface, 
the Sun rotates roughly once every month)
compared to the Sun's Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale becomes intractably large
compared to dynamical timescales \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are susceptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In other systems, 
the approximate state of the evolved solution can be estimated. There, a
set of initial conditions which is close to the evolved state can be 
derived analytically \cite{couston&all2017, brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
However, such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for hundreds of freefall timescales \cite{stevens&all2011}, much less the
thousands or millions freefall times required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of unequilibrated
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean vertical thermodynamic profile towards its evolved state. 
While a technique of this kind has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor do we find any study of its accuracy.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and the procedure for achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from evolution through a full thermal diffusion timescale. Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except where it is $\rho = \rho_0(1  - \alpha T_1)$ on the term where
the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, 
acts in the vertical momentum equation.
The equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0(z) + T_1(x, y, z, t)$ are the initial and fluctuating components of temperature, 
and $P$ is the kinematic pressure.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{R}$ and $\mathcal{P}$ 
are set by the Rayleigh (Ra) and Prandtl (Pr) numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr $= 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

In Eqns. (\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}),
linear terms are grouped on the left-hand side of the equations, while nonlinear terms
are found on the right-hand side. We timestep linear terms implicitly, and nonlinear
terms explicitly.
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns. (\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-stage 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.

Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally periodic with an extent of $x, y \in [0, \Gamma]$,
where $\Gamma = 2$ is the aspect ratio. 
In 2D simulations, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w = 0 \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. (\ref{eqn:bcs}) are the simplest
to implement in the process of accelerated evolution (section \ref{subsection:ae})
we study here.

The initial temperature profile is linearly unstable,
$T_0(z) = -z$. On top of this profile, we fill $T_1$ with
random white noise whose magnitude is $10^{-6}\mathcal{P}$.
This ensures that the initial perturbations are much smaller than the
evolved convective temperature perturbations, even at large Ra.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power.

@@

\subsection{The method of Accelerated Evolution}
\label{subsection:ae}
Here we describe a method of Accelerated Evolution (AE), which we use 
to rapidly evolve the thermodynamic state of convective simulations.  
We compare this AE method to Standard Evolution
(SE), in which we naively evolve the atmosphere for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. Both AE and SE simulatins begin with identical
initial conditions, as described in section \ref{sec:experiment}.
As Ra increases, and $\mathcal{P}$ decreases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
in simulation freefall time units (see table \ref{table:run_parameters}).

For an example of time saving achieved by using AE, we compare
energy traces at $S = 10^5$ from a SE run in Fig. \ref{fig:time_trace}a to an AE run
in Fig. \ref{fig:time_trace}c.
In Fig. \ref{fig:time_trace}a, the time evolution of the SE simulation is shown.
The solution grows exponentially from white noise during the
first $\sim 25$ $t_{\text{ff}}$. The solution then saturates and begins to slowly
equliibrate towards the proper isothermal profile in the interior of the domain.
We find that the mean atmospheric temperature and
kinetic energies are fully converged when $t = 4000$ $t_{\text{ff}}$ $= 0.35$ $t_{\kappa}$.
We show roughly the first thousand freefall time
units of evolution, as well as the evolved thermodynamic state reached after a full
thermal time of evolution.  In Fig. \ref{fig:time_trace}c, we show an AE solution
at the same parameters. The same linear growth phase occurs, but shortly after
the peak of convective transient we accelerate the convergence of the atmosphere
through the process which we describe below. We adjust the 1D vertical profile of
the atmosphere three times, as denoted by the three labeled
arrows in the graph numbered 1-3.  After the third profile adjustment, we find that
the atmosphere is nearly in its converged state and we can begin to sample the evolved
convective dynamics.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\kappa} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective enthalpy flux,
F$_{\text{E}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig. \ref{fig:time_trace}b.
At these early times, the flux profiles are highly asymmetric,
with more flux exiting the atmosphere at the upper boundary than 
the fixed-flux lower boundary can provide. As the atmosphere evolves towards
the isothermal profile specified by the upper (cold) boundary condition, excess
temperature throughout the atmosphere must leave the domain, and this appears in
the flux measurements in Fig. \ref{fig:time_trace}b where the flux exiting the upper
boundary is nearly 2000\% the flux entering the bottom of the domain. Once the atmospheric
temperature profile reaches its evolved state, the flux entering the bottom boundary
is equal to the flux exiting the upper boundary.  In general, this takes very long time
frames through SE (Fig. \ref{fig:time_trace}a), but AE (Fig. \ref{fig:time_trace}b)
can take a system whose fluxes are in a strongly disequilibrium state (Fig. \ref{fig:time_trace}c),
and quickly put them into a near-equilibrium state, as shown in Fig. \ref{fig:time_trace}d. In
this final state, there is no conductive flux in the interior, both boundaries carry the
same amount of flux, and convection carries that amount of flux through the interior.
The converged state achieved through AE is at most 5\% different from the SE solution, as shown in
Fig. \ref{fig:time_trace}e.

\begin{figure}[p!]
\includegraphics[width=\textwidth]{./figs/time_trace.png}
\caption{(a) Kinetic energy (black) and mean temperature (blue)  vs. time are shown
for a SE run at $S = 10^5$. The mean evolved values of kinetic energy and mean temperature,
averaged over the time shaded in green,
are denoted by the horizontal dashed lines. (b) The time- and horizontally-averaged
flux profiles are shown for the times highlighted in orange.
(c) The same quantities as in (a) are shown, but for AE at the same parameters.
The axes are scaled identically in (a) and (c), and the AE method is used three times, marked by
the numbered arrows in (c). The fluxes averaged over the green shaded region of (c)
are shown in (d). The difference between
the fluxes in the AE and SE solutions is shown in (e). \label{fig:time_trace} }
\end{figure}

In order to adjust the temperature profle to achieve AE, we calculate the total flux,
F$_{\text{tot}} =$ F$_{\text{E}} +$ F$_{\kappa}$, and then derive the profiles
\begin{equation}
f_{\text{E}}(z) = \frac{\text{F}_{\text{E}}}{\text{F}_{\text{tot}}},\qquad
f_{\kappa}(z) = \frac{\text{F}_{\kappa}}{\text{F}_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{E}}(z) = 1$
and $f_{\kappa}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\kappa}(z) = 1$ and $f_{\text{E}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, in order to reach the converged state, 
the flux through the atmosphere must be decreased by some amount,
\begin{equation}
\xi(z) \equiv \frac{\text{F}_{\text{B}}}{\text{F}_{\text{tot}}},
\label{eqn:xi}
\end{equation}
where $\text{F}_{\text{B}} = \mathcal{P}$ is the amount of flux that enters the
bottom of the atmosphere.

In a time-stationary state, the horizontal- and time-average of
Eqns. (\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}), neglecting terms which
vanish due to symmetry, are
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y, \text{ ev}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{E, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Here, we construct $\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}} =
\xi(z) \angles{\bm{u}\times\bm{\omega}}_{x, y}$ and $\text{F}_{\text{E, ev}} = \xi(z) \text{F}_{\text{E}}$
from our unevolved state, and solve for $\angles{\varpi}_{x,y}$ and $\angles{T_1}_{x,y}$.
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, 
statistically stationary state. Furthermore, under the specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$,
\emph{the mean thermodynamic structure of the system is fully specified}.
